|

### 2016-01-08 manuscript/thesis_pve.Rmd
figure double check
MT edits to thesis, made in all versions

pvetab1 <- "2.1"
pvetab2 <- "2.2"
pvetab3 <- "2.3"
pvetabS <- "2.S3"

@@

Interpreting limitations to plant growth requires understanding the balance between carbon (C) source and sink activity. This study used manipulations of soil volume to test how growth is coupled to physiology, C allocation, and sink activity in *Eucalyptus tereticornis* seedlings. We grew individual seedlings in a large range of container sizes and planted containers flush to the soil alongside naturally planted freely-rooted (free) seedlings. We developed a seedling growth model that utilized leaf photosynthesis rates (*A~n~*) to allocate daily C uptake towards mass growth of stems, leaves and roots. Reduced soil volume was expected to induce rapid negative effects on growth and physiology compared to free seedlings. It was hypothesized that the soil volume effect would be largest in the smallest containers, negatively impacting mass partitioning belowground. An accumulation of leaf non-structural carbohydrates, resulting from reduced belowground sink strength, was expected to correlate to reductions in photosynthetic capacity. We observed a negative effect of container volume on aboveground growth soon after the experiment started. Although growth was consistently different across soil volumes, dry mass partitioning to leaves, stems and roots was unchanged after 120 days. Photosynthetic capacity was significantly reduced in containers compared to free seedlings, and was related to both leaf nitrogen content and starch accumulation. We then asked whether the observed reductions in *A~n~* explained the observed differences in seedling biomass. We found that although belowground sink limitation resulted in down regulation of *A~n~*, these reductions were not large enough to explain observed growth responses. Thus, as *A~n~* and growth were not tightly coordinated, excess photosynthetic C not attributed to biomass resulted in seedlings with soil volume restriction. These results highlight the need to further utilize mass balance approaches when evaluating plant C allocation and confirm that important feedbacks exist between belowground sink strength and leaf C uptake.

@@

\
In woody species, the coordination of *A~n~* and growth has been studied with manipulations of C source activity. Examples included elevated CO~2~ experiments, for example FACE [reviewed in @ainsworth2005have], and partial defoliation experiments. Elevated CO~2~ has been shown to increase *A~n~* [@drake1997more; @ainsworth2007response] and across four FACE experiments this resulted in a stimulation of 23 % in forest biomass production [@norby2005forest]. Evidence from a wide range of elevated CO~2~ experiments, however, also reveals that even with an average photosynthetic enhancement of over 30 %, the biomass growth rate only increases by around 10 % [@kirschbaum2011does]. In partial defoliation experiments, increases in *A~n~* of the remaining foliage are commonly shown, yet are attributed to various mechanisms, including reduction in end product inhibition [@iglesias2002regulation; @zhou2003changes; @handa2005test], enhanced biochemical activity [@ovaska1993b;@layne1995end], increased stomatal conductance [@layne1995end], enhanced leaf nutrient status [@turnbull2007increased] and regulatory sugar signaling [@eyles2013whole]. However, increases in *A~n~* in defoliation experiments did not always produce increased growth due to reductions in meristem sink strength [@palacio2012fast], C limitation to mycorrhizal colonization [@markkola2004defoliation], or an overall decrease in whole plant C gain [@ovaska1993a]. These manipulations of C source activity expose unresolved issues with how changes in *A~n~* do not always infer proportional responses in growth.

@@

\
Six container volumes were used ranging from 5 L to 35 L, with a 22.5 cm diameter, and lengths ranging from 15 to 100 cm. Containers were constructed of PVC pipe and were filled with local top soil (described above). Soil in each container was packed to achieve a target soil bulk density that matched local soil conditions of 1.7 g cm^-3^. A Imidacloprid (BAYER CropScience) insecticide tablet was planted 5 cm below the roots of each seedling. Containers were planted flush with the soil surface inside metal sleeves, designed to minimize excess air space between the container and outside soil while also allowing for container removal. This allowed for soil temperatures in containers to reflect conditions of naturally planted (free) seedlings. Each experimental block (n=7) contained a complete replicate set of six container volumes as well as one free seedling, with 1 m^2^ spacing. For each free seedling, used as the control, a 1 m^2^ subplot was excavated to the hard layer and replaced with the same soil used in each container. A border of root exclusion material was buried 0.25 m deep and extended 0.25 m above the ground surface around each subplot to exclude local vegetation.

\
Plants were watered weekly or when needed to maintain soil moisture at field capacity (13-15 %). Drain systems were built into each pot to prevent pooling of water throughout the experiment. Pooling of water could lead to an anaerobic environment around the root that could hinder the uptake of water through reduced root conductance [@poorter2009causes], an undesired experimental artifact. A collection compartment in the bottom of containers, containing gravel covered by root exclusion mesh, was used to collect excess water for 20, 25, and 35 l containers. Plastic tubing (6 mm diameter) was inset into the gravel layer and extended through the top of the container. A lysimeter pump was then used to suction excess water, through the tubing, as needed. For small containers (5, 10, and 15 L) a simple bottom plug was used to drain excess water from the gravel compartment. Each containers was inspected after every rainfall event to determine if pooling had occurred.

@@

\
We then utilized this model to test the hypothesis that the effects of belowground sink limitation on rates of leaf *A~n~* where sufficient to accurately predict overall seedling biomass production after 120 days. Each model run utilized changes in *A~n~* and leaf mass fraction (LMF), with values of stem and root respiration rates, to generate total seedling mass and leaf area after 120 days.  Cumulative net leaf C gain for each treatment was equal to the sum of each value of C~day,i~ over 120 days and final seedling C was assumed to equal half of the final mass for both modeled and observed seedlings.

@@

##2.4.3 Gas exchange and photosynthetic parameters
At the first measurement campaign, both A~sat~ and A~max~ were significantly higher in the free seedling treatment compared to seedlings in containers (both P < 0.001). Across all measurement campaigns mean A~sat~ (Figure `r asatfig`) and A~max~ (Table `r pvetab3`) were consistently higher in free seedlings than in containers (26 % and 29 %, respectively). The relationships between photosynthetic capacity, leaf starch and leaf N on a mass basis was marginally significant (P = 0.058), but A~max~ on a mass basis was highly correlated to both leaf N content and leaf starch (both P < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationship of A~max~ to either leaf N content or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments A~max~ was higher when leaf N was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). A~max~  was also lower when leaf starch was high as higher leaf N often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, A~max~ was positively correlated with final harvest biomass across all seedlings (P < 0.001).
\
```{r asat, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE, message=FALSE, dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/A_stats_plotting.R")
```  
\
**Figure 2.4**. Soil volume treatment means &pm; standard error, across all measurement campaigns (n = 6), of light saturated rates of photosynthesis at 25&deg;C. Different letters represent significant differences between treatments.

@@

\
Here, reductions in *A~n~* were positively correlated with decreases in leaf N and leaf N was considerably reduced for seedlings in containers. As leaf N reductions were detected with TNC-free leaf mass, this suggests that physical root restriction or decreased supply likely affected seedling N uptake instead of TNC dilution in leaves in the smallest containers. Root N at the end of the experiment was on average higher in free seedlings but not consistently higher than every container volume treatment. Unrestricted mycorrhizal recruitment could have instead facilitated the increases in leaf N in free seedlings, but this effect is unknown. Combined with the fact that soil N declined evenly across all treatments, there was no clear mechanism present to identify changes in root N uptake between free seedlings and seedlings in containers. In these already low quality soils, it is possible that seedlings in containers simply grew into increasing N limitation which negatively affected belowground sink strength. Although no clear feedback could be determined between the available soil N pool and decreases in leaf N, the effects of belowground sink limitation on *A~n~* of seedlings in containers was evident throughout the experiment.

@@

\
The lack of detected shifts in partitioning to fine roots provides evidence against an optimal foraging strategy for seedlings in containers. This could be because lateral root development is affected by inanimate obstacles and avoiding growth towards container walls could improve the efficiency of resource allocation [@falik2005root]. The sensitivity of roots to their own exudates near obstructions may prevent further growth [@semchenko2008foraging]. Here, we found that FRLD was highest in smallest containers suggesting that root restriction likely occurred as simple function of available rooting space. Additionally, physical restriction of root proliferation could have impacted root development and morphology prior to shifts in mass partitioning. Here, increases in SRL where detected in several of the soil volume treatments. This is not surprising as plants in containers have been shown to have different root morphology to field grown plants [@nesmith1998effect]. The poor soil quality used in our experiment and root restriction, however, likely decreased the capacity of this morphological response to increase N uptake.

@@

\
Increasing rates of non-leaf respiration (M~S2~) improved biomass predictions but to a far lesser extent than changes to leaf C allocation. We also show that very large increases in non-leaf respiration rates would have been required to accurately predict observed seedling biomass. Although the fraction of photosynthate used in respiration is known to vary among species and is sensitive to changes in growth rates [@Lambers2008], results from M~S2~ highlight a lack of knowledge regarding how respiration rates of individual tissues, within a single plant, maybe be differentially affected by environmental change. This is noteworthy, as C balance is a delicate equilibrium between fluxes of *A~n~* and respiration, partial accounting of C dynamics can easily lead to erroneous conclusions [@valentini2000respiration]. These results infer that using fixed rates of respiration in models likely underestimates plant responses to sink limitation. Thus, we agree with Delucia et al. [-@delucia2007forest] that it is likely inappropriate to assume that respiration is a constant fraction of gross primary production in models. Our findings reveal that a combination of different mechanisms, beyond *A~n~*, are likely at play in driving the observed seedling biomass response to sink manipulation. However, the degree to which these mechanisms will regulate growth will undoubtedly shift across different experimental manipulations and plant species.

##2.5.4 Conclusions
With a novel field-based design we detected a massive effect of container volume on seedling growth, not between containers but with naturally planted seedlings. This is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles, but how these results actually reflect field-grown plants has seldom been studied. Although biomass partitioning was conserved, our empirical and model results suggest that the amount of photosynthate allocated to non-biomass pools such as TNC or respiration were likely altered by sink inhibition. The debate over how rates of photosynthesis affect plant growth or to what degree these rates are instead controlled by growth has existed for decades [@sweet1966role]. Our combined empirical and modeling approach shows that when non-photosynthetic parameters were kept constant changes in *A~n~* were not able to fully to predict changes in growth, an important distinction often missed in studies that manipulate source/sink activity. K&#246;rner [-@korner2013growth] suggests that it is the norm for sink activity to feedback onto source activity, causing growth to control *A~n~* through the demand for C. Although this may be true, our results infer that quantifying the fate of assimilated C into known pools of growth, storage and C loss are needed prior to addressing this debate. Our modelling results agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents. The approach used here has the flexibility to account for multiple drivers of C allocation and provides an avenue to address future questions regarding the impact of environmental change on plant growth.

@@

```{r model_scenarios, fig.width=6, fig.height=8, echo=FALSE, dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/massmodel_optimize_scenarios.R")
```  
\
\
\
**Figure 2.S1**. Sensitivity testing of seedling growth model to different carbon allocation strategies including; constraints of leaf mass fraction to treatment specific final harvest values (a) and increases in respiration of non-leaf tissue components by 50 % (b).  Open and filled symbols represent default model and harvest values, while shaded symbols represent model sensitivity to each scenario by soil volume treatment. Both seedling carbon mass and daily carbon assimilation were first scaled to the free seedling control.

>>>>

pvetab1 <- "2.1"
pvetab2 <- "2.2"
pvetab3 <- "2.3"
pvetabS <- "2.S1"

@@

Interpreting limitations to plant growth requires understanding the balance between carbon (C) source and sink activity. This study used manipulations of soil volume to test how growth is coupled to physiology, C allocation, and sink activity in *Eucalyptus tereticornis* seedlings. We grew individual seedlings in a large range of container sizes and planted containers flush to the soil alongside naturally planted freely-rooted (free) seedlings. We developed a seedling growth model that utilized leaf photosynthesis rates (*A~n~*) to allocate daily C uptake towards mass growth of stems, leaves and roots. Reduced soil volume was expected to induce rapid negative effects on growth and physiology compared to free seedlings. It was hypothesized that the soil volume effect would be largest in the smallest containers, negatively impacting mass partitioning belowground. An accumulation of leaf non-structural carbohydrates, resulting from reduced belowground sink strength, was expected to correlate to reductions in photosynthetic capacity. We observed a negative effect of container volume on aboveground growth soon after the experiment started. Although growth was consistently different across soil volumes, dry mass partitioning to leaves, stems and roots was unchanged after 120 days. Photosynthetic capacity was significantly reduced in containers compared to free seedlings, and was related to both reduced leaf nitrogen content and starch accumulation. We then asked whether the observed reductions in *A~n~* explained the observed differences in seedling biomass. We found that although belowground sink limitation resulted in down regulation of *A~n~*, these reductions were not large enough to explain observed growth responses. Thus, as *A~n~* and growth were not tightly coordinated, excess photosynthetic C not attributed to biomass resulted in seedlings with soil volume restriction. These results highlight the need to further utilize mass balance approaches when evaluating plant C allocation and confirm that important feedbacks exist between belowground sink strength and leaf C uptake.

@@

\
In woody species, the coordination of *A~n~* and growth has been studied with manipulations of C source activity. Examples included elevated CO~2~ experiments, for example FACE [reviewed in @ainsworth2005have], and partial defoliation experiments. Elevated CO~2~ has been shown to increase *A~n~* [@drake1997more; @ainsworth2007response] and across four FACE experiments this resulted in a stimulation of 23 % in forest biomass production [@norby2005forest]. Evidence from a wide range of elevated CO~2~ experiments, however, also reveals that even with an average photosynthetic enhancement of over 30 %, the biomass growth rate only increases by around 10 % [@kirschbaum2011does]. In partial defoliation experiments, increases in *A~n~* of the remaining foliage are commonly shown, yet are attributed to various mechanisms, including reduction in end-product inhibition [@iglesias2002regulation; @zhou2003changes; @handa2005test], enhanced biochemical activity [@ovaska1993b;@layne1995end], increased stomatal conductance [@layne1995end], enhanced leaf nutrient status [@turnbull2007increased] and regulatory sugar signaling [@eyles2013whole]. However, increases in *A~n~* in defoliation experiments did not always produce increased growth due to reductions in meristem sink strength [@palacio2012fast], C limitation to mycorrhizal colonization [@markkola2004defoliation], or an overall decrease in whole plant C gain [@ovaska1993a]. These manipulations of C source activity expose unresolved issues with how changes in *A~n~* do not always correspond with proportional responses in growth.

@@

\
Six container volumes were used ranging from 5 L to 35 L, with a 22.5 cm diameter, and lengths ranging from 15 to 100 cm. Containers were constructed of PVC pipe and were filled with local top soil (described above). Soil in each container was packed to achieve a target soil bulk density that matched local soil conditions of 1.7 g cm^-3^. A Imidacloprid (BAYER CropScience) insecticide tablet was planted 5 cm below the roots of each seedling. Containers were planted flush with the soil surface inside metal sleeves, designed to minimize excess air space between the container and outside soil while also allowing for container removal. This allowed for soil temperatures in containers to reflect conditions of naturally planted (free) seedlings. Each experimental block (n=7) contained a complete replicate set of six container volumes as well as one free seedling, with 1 m^2^ spacing. For each free seedling, used as the control, a 1 m^2^ subplot was excavated to the hard layer and replaced with the same soil used in each container. A border of root exclusion material was buried 0.25 m deep and extended 0.25 m above the ground surface around each subplot to exclude local vegetation, which was further kept out by periodic weeding.

\
Plants were watered weekly or when needed to maintain soil moisture at field capacity (13-15 %). Drain systems were built into each pot to prevent pooling of water throughout the experiment. Pooling of water could lead to an anaerobic environment around the root that could hinder the uptake of water through reduced root conductance [@poorter2009causes], an undesired experimental artifact. A collection compartment in the bottom of containers, containing gravel covered by root exclusion mesh, was used to collect excess water for 20, 25, and 35 l containers. Plastic tubing (6 mm diameter) was inset into the gravel layer and extended through the top of the container. A lysimeter pump was then used to suction excess water, through the tubing, as needed. For small containers (5, 10, and 15 L) a simple bottom plug was used to drain excess water from the gravel compartment. Each container was inspected after every watering and rainfall event to determine if pooling had occurred.

@@

\
We then utilized this model to test the hypothesis that the effects of belowground sink limitation on rates of leaf *A~n~* were sufficient to accurately predict overall seedling biomass production after 120 days. Each model run utilized changes in *A~n~* and leaf mass fraction (LMF), with values of stem and root respiration rates, to generate total seedling mass and leaf area after 120 days.  Cumulative net leaf C gain for each treatment was equal to the sum of each value of C~day,i~ over 120 days and final seedling C was assumed to equal half of the final dry mass for both modeled and observed seedlings.

@@

##2.4.3 Gas exchange and photosynthetic parameters
At the first measurement campaign, both A~sat~ and A~max~ were significantly higher in the free seedling treatment compared to seedlings in containers (both P < 0.001). Across all measurement campaigns mean A~sat~ (Figure `r asatfig`) and A~max~ (Table `r pvetab3`) were consistently higher in free seedlings than in containers (26 % and 29 %, respectively). The relationships between photosynthetic capacity, leaf starch and leaf N on a mass basis was marginally significant (P = 0.058), but A~max~ on a mass basis was highly correlated to both leaf N content and leaf starch (both P < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationship of A~max~ to either leaf N content or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments A~max~ was higher when leaf N was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). A~max~  was also lower when leaf starch was high as higher leaf N often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, A~max~ was positively correlated with final harvest biomass across all seedlings (P < 0.001).
\
```{r asat, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE, message=FALSE, dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/A_stats_plotting.R")
```  
\
**Figure 2.4**. Soil volume treatment means &pm; standard error, across all measurement campaigns (n = 6), of light saturated rates of photosynthesis at 25 &deg;C. Different letters represent significant differences between treatments.

@@

\
Here, reductions in *A~n~* were positively correlated with decreases in leaf N and leaf N was considerably reduced for seedlings in containers. As leaf N reductions were detected with TNC-free leaf mass, this suggests that physical root restriction or decreased supply likely affected seedling N uptake instead of TNC dilution in leaves in the smallest containers. Root N at the end of the experiment was on average higher in free seedlings but not consistently higher than every container volume treatment. Unrestricted mycorrhizal recruitment could have instead facilitated the increases in leaf N in free seedlings, but this effect is unknown. Combined with the fact that soil N declined evenly across all treatments, there was no clear mechanism present to identify changes in root N uptake between free seedlings and seedlings in containers. In these already low fertility soils, it is possible that seedlings in containers simply grew into increasing N limitation which negatively affected belowground sink strength. Although no clear feedback could be determined between the available soil N pool and decreases in leaf N, the effects of belowground sink limitation on *A~n~* of seedlings in containers was evident throughout the experiment.

@@

\
The lack of detected shifts in partitioning to fine roots provides evidence against an optimal foraging strategy for seedlings in containers. This could be because lateral root development is affected by inanimate obstacles and avoiding growth towards container walls could improve the efficiency of resource allocation [@falik2005root]. The sensitivity of roots to their own exudates near obstructions may prevent further growth [@semchenko2008foraging]. Here, we found that FRLD was highest in smallest containers suggesting that root restriction likely occurred as simple function of available rooting space. Additionally, physical restriction of root proliferation could have impacted root development and morphology prior to shifts in mass partitioning. Here, increases in SRL were detected in several of the soil volume treatments. This is not surprising as plants in containers have been shown to have different root morphology to field grown plants [@nesmith1998effect]. The poor soil quality used in our experiment and root restriction, however, likely decreased the capacity of this morphological response to increase N uptake.

@@

\
Increasing rates of non-leaf respiration (M~S2~) improved biomass predictions but to a far lesser extent than changes to leaf C allocation. We also show that very large increases in non-leaf respiration rates would have been required to accurately predict observed seedling biomass. Although the fraction of photosynthate used in respiration is known to vary among species and is sensitive to changes in growth rates [@Lambers2008], results from M~S2~ highlight a lack of knowledge regarding how respiration rates of individual tissues, within a single plant, maybe be differentially affected by environmental change. This is noteworthy, as C balance is a delicate equilibrium between fluxes of *A~n~* and respiration, partial accounting of C dynamics can easily lead to erroneous conclusions [@valentini2000respiration]. These results imply that using fixed rates of respiration in models likely underestimates plant responses to sink limitation. Thus, we agree with Delucia et al. [-@delucia2007forest] that it is likely inappropriate to assume that respiration is a constant fraction of gross primary production in models. Our findings reveal that a combination of different mechanisms, beyond *A~n~*, are likely at play in driving the observed seedling biomass response to sink manipulation. However, the degree to which these mechanisms will regulate growth will undoubtedly shift across different experimental manipulations and plant species.

##2.5.4 Conclusions
With a novel field-based design we detected a massive effect of container volume on seedling growth, when compared with naturally planted seedlings. This is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles, but how these results actually reflect field-grown plants has seldom been studied. Although biomass partitioning was conserved, our empirical and model results suggest that the amount of photosynthate allocated to non-biomass pools such as TNC or respiration were likely altered by sink inhibition. The debate over how rates of photosynthesis affect plant growth or to what degree these rates are instead controlled by growth has existed for decades [@sweet1966role]. Our combined empirical and modeling approach shows that when non-photosynthetic parameters were kept constant, changes in *A~n~* were not able to fully to predict changes in growth, an important distinction often missed in studies that manipulate source/sink activity. K&#246;rner [-@korner2013growth] suggests that it is the norm for sink activity to feedback onto source activity, causing growth to control *A~n~* through the demand for C. Although this may be true, our results imply that quantifying the fate of assimilated C into known pools of growth, storage and C loss are needed prior to addressing this debate. Our modelling results agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents. The approach used here has the flexibility to account for multiple drivers of C allocation and provides an avenue to address future questions regarding the impact of environmental change on plant growth.

@@

```{r model_scenarios, fig.width=6, fig.height=8, echo=FALSE, dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/massmodel_optimize_scenarios.R")
```  
\
\
\
**Figure 2.S1**. Sensitivity testing of the seedling growth model to different carbon allocation strategies including; constraints of leaf mass fraction to treatment specific final harvest values (a) and increases in respiration of non-leaf tissue components by 50 % (b).  Open and filled symbols represent default model and harvest values, while shaded symbols represent model sensitivity to each scenario by soil volume treatment. Both seedling carbon mass and daily carbon assimilation were first scaled to the free seedling control.
